---
title: "Statistical Inference Course Project Part 1"
author: "Ziyao Gao"
date: "10/30/2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment Description
Investigate the exponential distribution in R and compare it with the Central Limit Theorem.The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations.

Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials. You should:

1. Show the sample mean and compare it to the theoretical mean of the distribution.
2. Show how variable the sample is (via variance) and compare it to the theoretical      variance of the distribution.
3. Show that the distribution is approximately normal.

Enviroment requirement for this report:

```{r}
# install the pacakes needed 
library(knitr)
library(ggplot2)
# set seed to make the report reproducible
set.seed(12345)
```

# Simulation Exercise
In this project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations.

```{r}
# set variables for the simulation
n <- 40
lambda <- 0.2
simulations <- 1000

# Create a matrix of 1000 rows with the columns for random simulation 40 times
matrix <- matrix(rexp(simulations * n, rate = lambda), simulations, n)

# calcualte means and plot it through histogram 
sim_mean <- rowMeans(matrix)
hist(sim_mean, xlab = "Mean of 40 Exponentials", ylab = "Frequency",
     main = "Histrogram of Simulation Mean", col = "green")
```

# Sample Mean vs. Theoretical Mean
```{r}
sample_mean <- mean(sim_mean)
sample_mean

theoretical_mean <- 1/lambda
theoretical_mean
```
As we can see above, sample mean is really close to theoretical mean.

# Sample Variance vs. Theoretical Variance
```{r}
sample_var <- var(sim_mean)
sample_var

theoretical_var <- (1/lambda)^2/n
theoretical_var
```
As we can see above, sample variance is also very close to theoretical variance.

# Approximate normal distribution
First, create an approximate normal distribution and see how the sample aligns with it 
```{r}
plotdata <- data.frame(sim_mean)
g <- ggplot(plotdata, aes(x = sim_mean))
g = g + geom_histogram(aes(y = ..density..), colour = "blue", fill = "green")
g = g + geom_density(colour = "black", size = 1)
g = g + ggtitle("Histogram of Simulation Mean")
g
```

The plot above indicates that the histogram can be adeqautely approximated with the normal distribution.

Second, let's compare their confidence intervals as well:
```{r}
sample_conf_interval <- round(mean(sim_mean) + c(-1,1)*1.96*sd(sim_mean)/sqrt(n), 3)
sample_conf_interval

theoretical_conf_interval <- round(mean(theoretical_mean) + c(-1, 1)*1.96*sqrt(theoretical_var)/sqrt(n), 3)
theoretical_conf_interval
```
From the results we can see, the sample 95% confidence interval [4.729, 5.215] is close to theoretical confidence interval [4.755, 5.245] as well 

Third, plot Q-Q for quantitles.
```{r}
qqnorm(sim_mean, main = "Normal Q-Q Plot", xlab = "Theoretical Quantiles", ylab = "Sample Quantitles")
qqline(sim_mean, col = "red")
```

As we can see, the distribuion is approximately normal. 











